COVID-19 Vaccine Acceptance and Hesitancy in Low and Middle Income Countries: Replication Code

### 2021-06-22 4_paper_supplementary_files/temp_download.Rmd
update

@@

```{r, include=FALSE} 
empty = " " 
```
`r empty`

@@

We analyze COVID-19 vaccine acceptance across 15 survey samples covering ten low- and middle-income countries (LMICs) in Asia, Africa, and South America, Russia (an upper-middle-income country), and the United States, using survey responses from  `r nice(nrow(df))` individuals. We find considerably higher willingness to take a COVID-19 vaccine in LMIC samples (80% on average) compared to the United States (65%) and Russia (30%). Vaccine acceptance was primarily explained by an interest in personal protection against COVID-19, while concern about side effects was the most commonly expressed reason for reluctance. Health workers were the most trusted sources of information about COVID-19 vaccines. Our findings suggest that prioritizing vaccine distribution to LMICs should yield high returns in promoting global immunization coverage, and that vaccination campaigns in these countries should focus on translating acceptance into uptake. Messaging highlighting vaccine efficacy and safety, delivered by healthcare workers, may be most effective in addressing remaining hesitancy.

@@

A safe and effective vaccine against COVID-19 is a critical tool to control the pandemic. As of April 27, 2021, 22 vaccines had advanced to Stage 3 clinical trials [@who1] and more than a dozen vaccines had been approved in multiple countries[@track]. For example, as of April 27, 2021, Comirnaty (BioNTech/Pfizer) had been approved in  more than 80 countries, while the AstraZeneca vaccine had the most country authorizations at 92. [@track] At present, however, global vaccine distribution remains highly unequal, with much of the current supply directed toward high-income countries [@wouters2021challenges].

While effective and equitable distribution of COVID-19 vaccines is a key policy priority, ensuring the population’s acceptance is equally important. Trust in vaccines and the institutions that administer them are key determinants of the success of any vaccination campaign [@defigueiredo2020lancet]. Several studies have investigated willingness to take a potential COVID-19 vaccine in high-income countries [@boyon2020ipsos; @Malik2020; @kreps2020factors; @shekhar2021covid; @dror2020vaccine; @fisher2020attitudes], and some studies have included middle-income countries [@lazarus2020nature; @wouters2021challenges]. Less is known, however, about vaccine acceptance in low-income countries where large-scale vaccination  has yet to begin. Understanding the drivers of COVID-19 vaccine acceptance is of global concern, since a lag in vaccination in any country may result in the emergence and spread of new variants that can overcome vaccine- and prior disease-conferred immunity [@ong2021lack; @lancettask].

Our study complements the emerging global picture of COVID-19 vaccine acceptance by focusing primarily on lower income countries. We construct a sample of low- and middle-income countries (LMICs) with wide geographic coverage across Africa, Asia and Latin America. We move beyond documenting vaccine acceptance rates and analyze data on the reasons for acceptance and hesitancy, which is critical for informing the design of effective vaccine distribution and messaging.

Acceptance of childhood vaccination for common diseases---such as measles (MCV), Bacille Calmette-Guérin (BCG) and diphtheria, tetanus, and pertussis (DTP)---is generally high in LMICs, providing cause for optimism about the prospects for COVID-19 vaccine uptake. Table \ref{tab:otherv} summarizes general vaccine acceptance and coverage rates of childhood vaccines in 2018, prior to the current pandemic,  for the countries included in our study. Agreement on the importance of childhood vaccinations is markedly higher in these LMICs compared to Russia and the United States. Still, existing studies on COVID-19 vaccine acceptance document substantial variation, both across and within countries, including in settings with high background levels of other vaccinations [@defigueiredo2020lancet; @lazarus2020nature; @wouters2021challenges].

This literature cites concerns about COVID-19 vaccine safety, including concerns about the rapid pace of vaccine development, as reasons for hesitancy in higher-income settings [@boyon2020ipsos; @wouters2021challenges], but other reasons  may feature more prominently in LMICs. For instance, reported COVID-19 mortality rates have been consistently lower in LMICs relative to higher income countries [@mukherjee]. If individuals feel the risk of disease is less serious, they may be less willing to accept any perceived risks of vaccination [@brewer2007meta]. Previous studies of healthcare utilization in LMICs have also highlighted factors such as healthcare quality concerns [@christensen2020building], negative historical experiences involving foreign actors  [@Lowes2018; @martinez2021vaccines], weak support from traditional leaders [@Jegede2007], and mistrust in government [@BLAIR201789] as barriers to uptake, which could apply to COVID-19 vaccination as well.

To promote vaccination against COVID-19 and develop effective messaging strategies, we need to know whether people are willing to take COVID-19 vaccines, the reasons why they are willing or unwilling to do so, and the most trusted  sources in their decision-making. Our study investigates these questions using a common set of survey items deployed across 13 studies in Africa, South Asia, and Latin America: seven in low-income countries (Burkina Faso, Mozambique, Rwanda, Sierra Leone, Uganda), five in lower-middle-income countries (India, Nepal, Nigeria, Pakistan), and one in an upper-middle-income country (Colombia). We compare these findings to those from two countries at the forefront of vaccine research and development, Russia and the United States. 
To select studies to include in our sample, we conducted an internal search within Innovations for Poverty Action (IPA), the International Growth Center (IGC), and the Berlin Social Science Center (WZB) for projects with plans to collect survey data in the second half of 2020. Study PIs agreed to include a set of common questions about COVID-19 vaccine attitudes. This strategy was guided by the need to collect information quickly and cost effectively using a survey modality (phone) that was both safe given pandemic conditions and appropriate for contexts with limited internet coverage. The final set of samples included in our study therefore reflects populations that fall under the current research priorities at IGC, IPA and WZB and, in the case of IPA and IGC, donors that prioritize working in the Global South.

@@

Our main results are shown in Figure \@ref(fig:mainfigure1) and reproduced as [Appendix A](#appendixd) Table \ref{tab:maintabledis}. The first column provides overall acceptance rates in each study, while the remaining columns disaggregate acceptance rates by respondent characteristics. The “All LMICs” row reports averages for LMIC countries and excludes Russia and the USA.

We document meaningful variation in vaccine acceptance across and within LMICs, but generally high levels of acceptance in LMICs overall. The average acceptance rate across studies was `r ans_mean`% (95% CI `r ans_mean_low`--`r ans_mean_high`%), with a median of `r ans_median`, a range of `r ans_max - ans_min` percent points and an interquartile range of `r ans_iqr`. Our estimate of $\tau^2$ is `r  round(mp_tau, 3)` which implies a standard deviation over country averages of  `r round(mp_tau^.5, 3)`.

The acceptance rate in every LMIC sample was higher than in the USA (`r usa_ans %>% pull(estimate)`%, `r usa_ans %>% pull(conf.low)`--`r usa_ans %>% pull(conf.high)`%) and Russia (`r rus_ans %>% pull(estimate)`%, `r rus_ans %>% pull(conf.low)`--`r rus_ans %>% pull(conf.high)`%). Reported acceptance was lowest in `r bottom_all$group[1]` (`r bottom_all$estimate[1]`% `r bottom_all$conf.low[1]`--`r bottom_all$conf.high[1]`%) and `r bottom_all$group[2]`  (`r bottom_all$estimate[2]`%, `r bottom_all$conf.low[2]`--`r bottom_all$conf.high[2]`%). The case of Pakistan may be linked to negative historical experiences with foreign-led vaccination campaigns [@ali2019polio; @robbins2012cia]. This hesitancy may be particularly problematic, given the magnitude of the second wave in neighboring India just as vaccines are becoming available, with accelerating cases in Bangladesh, Pakistan and India threatening to overwhelm health infrastructure.

We find limited evidence of variation across demographic subgroups in our LMIC samples, as shown in [Appendix A](#appendixd) Table \ref{tab:dmeans}. Women were generally less willing to accept the vaccine (average difference about `r dif_gender` points, significant at $p<.01$). Respondents under age 24 and less educated respondents were marginally more willing to take the vaccine, but these differences were not statistically significant. The acceptance advantage among the less educated was most pronounced in Rwanda, Burkina Faso, and Uganda, and stands in contrast to the USA, where more educated respondents were significantly more willing to be vaccinated.

To better understand the reasoning behind vaccine acceptance, we asked those who were willing to take the vaccine why they would take it. We summarize these results in Table \ref{tab:yes}, with additional details in [Appendix A](#appendixd) Table \ref{tab:yesall}. The reason most commonly given for vaccine acceptance across samples was personal protection against COVID-19 infection. The average across the LMIC samples was `r yes_all$estimate`% (`r yes_all$conf.low`--`r yes_all$conf.high`%). In every individual study, this ranked as the most popular reason. In distant second place among LMIC respondents was family protection with an average of `r yes_all_2$estimate`% (`r yes_all_2$conf.low`--`r yes_all_2$conf.high`%). In comparison to protecting oneself and one’s family,  protecting one’s community did not feature prominently among stated reasons for acceptance. These reasons do not vary substantially by age group, as shown in Table \ref{tab:countrydiff}. Self-protection also ranked as the most common reason for taking the vaccine in Russia (`r yes_rus$estimate`%, `r yes_rus$conf.low`--`r yes_rus$conf.high`%) and the USA (`r yes_usa$estimate`%, `r yes_usa$conf.low`--`r yes_usa$conf.high`%).

Figure \@ref(fig:fig2paper) summarizes the reasons given by respondents who said they were not willing to take a COVID-19 vaccine. Concern about side effects was the most frequently expressed reason for reluctance in our  LMIC samples. This concern was particularly evident among African country samples. In studies  `r side_top$group[1]` (`r side_top$estimate[1]`% `r side_top$conf.low[1]`--`r side_top$conf.high[1]`%), `r side_top$group[2]`  (`r side_top$estimate[2]`%, `r side_top$conf.low[2]`--`r side_top$conf.high[2]`%), `r side_top$group[3]` (`r side_top$estimate[3]`% `r side_top$conf.low[3]`--`r side_top$conf.high[3]`%) and `r side_top$group[4]` (`r side_top$estimate[4]`% `r side_top$conf.low[4]`--`r side_top$conf.high[4]`%), more than half of respondents unwilling to take the vaccine cited worries about side effects.  Respondents in Russia (`r side_rus$estimate`%, `r side_rus$conf.low`--`r side_rus$conf.high`%) and even more in the USA (`r side_usa$estimate`%, `r side_usa$conf.low`--`r side_usa$conf.high`%), frequently reported this same concern.

Study samples `r effective_top$group[1]` (`r effective_top$estimate[1]`%, `r effective_top$conf.low[1]`--`r effective_top$conf.high[1]`%), `r effective_top$group[2]`  (`r effective_top$estimate[2]`%, `r effective_top$conf.low[2]`--`r effective_top$conf.high[2]`%) and `r effective_top$group[3]` (`r effective_top$estimate[3]`%, `r effective_top$conf.low[3]`--`r effective_top$conf.high[3]`%) showed relatively high levels of skepticism about vaccine effectiveness among hesitant respondents. This was also true in Russia (`r effective_rus$estimate`%, `r effective_rus$conf.low`--`r effective_rus$conf.high`%) and the USA (`r effective_usa$estimate`%, `r effective_usa$conf.low`--`r effective_usa$conf.high`%). In addition, hesitant respondents listed lack of concern about COVID-19 infection as a reason not to be vaccinated. Study samples `r ill_top$group[1]` (`r ill_top$estimate[1]`% `r ill_top$conf.low[1]`--`r ill_top$conf.high[1]`%), `r ill_top$group[2]`  (`r ill_top$estimate[2]`%, `r ill_top$conf.low[2]`--`r ill_top$conf.high[2]`%) and `r ill_top$group[3]` (`r ill_top$estimate[3]`% `r ill_top$conf.low[3]`--`r ill_top$conf.high[3]`%) reported high rates of this answer among hesitant respondents.

In Figure \@ref(fig:fig3paper) we report respondents’ most trusted source of guidance when deciding whether to take a COVID-19 vaccine. Results from Figure \@ref(fig:fig3paper) are reproduced as  [Appendix A](#appendixd) Table \ref{tab:trust}.  [Appendix B](#appendixb) Table \ref{tab:q4} presents a complete description of response recoding from individual studies.

We find striking consistency in most trusted sources across studies. In all but one sample, including those from Russia and the USA, respondents identified the health system as the most trustworthy source to help them decide whether to take the COVID-19 vaccine. The average across LMICs was `r trust_all$estimate`% (`r trust_all$conf.low`--`r trust_all$conf.high`%). Respondents in `r trust_top$group[1]` (`r trust_top$estimate[1]`%, `r trust_top$conf.low[1]`--`r trust_top$conf.high[1]`%), `r trust_top$group[2]`  (`r trust_top$estimate[2]`%, `r trust_top$conf.low[2]`--`r trust_top$conf.high[2]`%) and `r trust_top$group[3]` (`r trust_top$estimate[3]`%, `r trust_top$conf.low[3]`--`r trust_top$conf.high[3]`%) cited  health workers most often. Sierra Leone had the highest trust in health workers and the Ministry of Health, potentially due to investments in public health following the 2014-2015 Ebola epidemic [@deserrano].

The next most cited sources were family and/or friends in  `r trust_fam$group[1]` (`r trust_fam$estimate[1]`%, `r trust_fam$conf.low[1]`--`r trust_fam$conf.high[1]`%), `r trust_fam$group[2]`  (`r trust_fam$estimate[2]`%, `r trust_fam$conf.low[2]`--`r trust_fam$conf.high[2]`%), `r trust_fam$group[3]` (`r trust_fam$estimate[3]`%, `r trust_fam$conf.low[3]`--`r trust_fam$conf.high[3]`%) and `r trust_fam$group[4]` (`r trust_fam$estimate[4]`%, `r trust_fam$conf.low[4]`--`r trust_fam$conf.high[4]`%). Across the pooled samples, women were 3 percentage points more likely to rely on family and friends than male respondents, though this difference was not statistically significant (Figure \@ref(fig:genderhist) in [Appendix D](#appendixd)).

By contrast, in `r trust_gov$group[1]` (`r trust_gov$estimate[1]`%, `r trust_gov$conf.low[1]`--`r trust_gov$conf.high[1]`%), `r trust_gov$group[2]`  (`r trust_gov$estimate[2]`%, `r trust_gov$conf.low[2]`--`r trust_gov$conf.high[2]`%), `r trust_gov$group[3]` (`r trust_gov$estimate[3]`%, `r trust_gov$conf.low[3]`--`r trust_gov$conf.high[3]`%) and `r trust_gov$group[4]` (`r trust_gov$estimate[4]`%, `r trust_gov$conf.low[4]`--`r trust_gov$conf.high[4]`%), government was the second most frequently cited. Religious leaders and celebrities were not seen as the top sources of guidance by many respondents in any sample other than Nepal, where many respondents most trusted famous people (`r trust_npl$estimate`%, `r trust_npl$conf.low`--`r trust_npl$conf.high`%).

>>>>

Julio S. Solis Arce,^[WZB Berlin Social Science Center, $^2$Innovations for Poverty Action (IPA), $^3$International Growth Centre (IGC), $^4$Wageningen University & Research, $^5$International Center for the Study of Institutions and Development (HSE University, Moscow, Russia), $^6$Columbia University, $^7$Yale Institute for Global Health, $^8$Busara Center for Behavioral Economics, $^{10}$Department of Sociology, University of Lagos, $^{11}$Busara Nigeria, $^{12}$Agricultural and Rural Development Secretariat, Federal Capital Territory Administration (Abuja, Nigeria), $^{13}$Nova School of Business and Economics, $^{14}$The Institute for Fiscal Studies, $^{15}$Lahore University of Management Sciences, $^{16}$Innovations for Poverty Action (IPA) Uganda, $^{17}$Morsel Research & Development, $^{18}$University of St. Andrews,  $^{19}$Redes Peru, $^{20}$Stockholm School of Economics and Misum, $^{21}$Ghent University, Department of Economics, $^{22}$Innovations for Poverty Action (IPA) Colombia, $^{23}$Institute of Development and Economic Alternatives, $^{24}$Innovations for Poverty Action (IPA) Sierra Leone,$^{25}$NOVAFRICA, $^{26}$Trinity College Dublin, $^{27}$Institute of Public Administration and Management, University of Sierra Leone, $^{28}$Centre for the Study of Labour and Mobility (CESLAM), $^{29}$Cornell University, $^{30}$University of Illinois Chicago, $^{31}$Innovations for Poverty Action (IPA) Rwanda, $^{32}$Associação NOVAFRICA para o Desenvolvimento Empresarial e Económico de Moçambique, $^{33}$Innovations for Poverty Action (IPA) Burkina Faso, $^{34}$NYU Abu Dhabi, $^{35}$Centre for Economic Research in Pakistan (CERP), $^{36}$Yale Research Initiative on Innovation and Scale (Y-RISE), $^{37}$Princeton University, $^{38}$Institute for International Economic Studies (IIES), Stockholm University, $^{39}$Tufts University, $^{40}$University of Michigan, $^{41}$Kellogg School of Management at Northwestern University, $^{42}$London School of Economics, $^{43}$Yale University] Shana S. Warren,$^2$ Niccolò F. Meriggi,$^3$ Alexandra Scacco,$^1$ Nina McMurry,$^1$ Maarten Voors,$^4$ Georgiy Syunyaev,$^{1,5,6}$, Amyn Abdul Malik,$^7$ Samya Aboutajdine,$^3$ Opeyemi Adeojo,$^{8, 10}$ Deborah Anigo,$^{11, 12}$ Alex Armand,$^{13, 14}$ Saher Asad,$^{15}$ Martin Atyera,$^{16}$ Britta Augsburg,$^{14}$ Manisha Awasthi,$^{17}$ Gloria Eden Ayesiga,$^{16}$ Antonella Bancalari,$^{14, 18, 19}$ Martina Björkman Nyqvist,$^{20}$ Ekaterina Borisova,$^{5, 21}$ Constantin Manuel Bosancianu,$^1$ Magarita Rosa Cabra García,$^22$ Ali Cheema,$^{15, 23}$ Elliott Collins,$^{2}$ Filippo Cuccaro,$^{24}$ Ahsan Zia Farooqi,$^{23}$ Tatheer Fatima,$^{17}$ Mattia Fracchia,$^{13, 25}$ Mery Len Galindo Soria,$^{22}$ Andrea Guariso,$^{26}$ Ali Hasanain,$^{15}$ Sofía Jaramillo Arciniegas,$^{22}$ Sellu Kallon,$^{4, 27}$ Anthony Kamwesigye,$^{16}$ Arjun Kharel,$^{28}$ Sarah Kreps,$^{29}$ Madison Levine,$^{4}$ Rebecca Littman,$^{30}$ Shan-e-Mohammad Malik,$^{23}$ Gisele Manirabaruta,$^{31}$ Jean Léodomir Habarimana Mfura,$^{31}$ Fatoma Momoh,$^{24}$ Alberto Mucauque,$^{32}$ Imamo Mussa,$^{32}$ Jean Aime Nsabimana,$^{31}$ Isaac Obara,$^{8}$ María Juliana Otálora,$^{22}$ Béchir Wendemi Ouédraogo,$^{33}$ Touba Bakary Pare,$^{33}$ Melina R. Platas,$^{34}$ Laura Polanco,$^{22}$ Javaeria Ashraf Qureshi,$^{30}$ Mariam Raheem,$^{35}$ Vasudha Ramakrishna,$^{36}$ Ismail Rendrá,$^{32}$ Taimur Shah,$^{35}$ Sarene Eyla Shaked,$^{16}$ Jacob N. Shapiro, $^{37}$ Jakob Svensson,$^{38}$ Ahsan Tariq,$^{23}$ Achille Mignondo Tchibozo,$^{33}$ Hamid Ali Tiwana,$^{23}$ Bhartendu Trivedi,$^{17}$ Corey Vernot, $^{36}$ Pedro C. Vicente,$^{13, 25}$ Laurin B Weissinger,$^{39}$ Basit Zafar,$^{35, 40}$ Baobao Zhang,$^{29}$ Dean Karlan,$^{2, 41}$ Michael Callen,$^{42}$ Matthieu Teachout,$^{3}$ Macartan Humphreys,$^{1, 6}$ Ahmed Mushfiq Mobarak,$^{43}$ Saad B. Omer$^{7}$.

@@

```{r} 
empty = " " 
```
`r empty`

@@

We analyze COVID-19 vaccine acceptance across 15 survey samples covering ten low- and middle-income countries (LMICs) in Asia, Africa, and South America, Russia (an upper-middle-income country), and the United States, including  `r nice(nrow(df))` individuals. We find considerably higher willingness to take a COVID-19 vaccine in our LMIC samples (mean `r ans_mean`%; median `r ans_median`%; range of `r ans_max - ans_min`%) compared to the United States (mean `r usa_ans %>% pull(estimate)`%) and Russia (mean `r rus_ans %>% pull(estimate)`%). Vaccine acceptance is primarily explained by an interest in personal protection against COVID-19, while concern about side effects is the most common reason for hesitancy. Health workers are the most trusted sources of guidance about COVID-19 vaccines. Evidence from this sample of LMICs suggests that prioritizing vaccine distribution to the Global South should yield high returns in advancing global immunization coverage. Vaccination campaigns should focus on translating the high levels of stated acceptance into actual uptake. Messages highlighting vaccine efficacy and safety, delivered by healthcare workers, may be most effective for addressing remaining hesitancy.

@@

A safe and effective vaccine is a critical tool to control the COVID-19 pandemic. As of May 27, 2021, 23 vaccines had advanced to Stage 3 clinical trials [@who1] and more than a dozen had been approved in multiple countries[@track]. The BioNTech/Pfizer (Comirnaty) vaccine, for example, has been approved in more than 80 countries, while the AstraZeneca vaccine has the most country authorizations at 98 [@track]. At present, however, global vaccine distribution remains highly unequal, with much of the current supply directed toward high-income countries [@wouters2021challenges].

While effective and equitable distribution of COVID-19 vaccines is a key policy priority, ensuring acceptance is just as important. Trust in vaccines and the institutions that administer them are key determinants of the success of any vaccination campaign [@defigueiredo2020lancet]. Several studies have investigated willingness to take a potential COVID-19 vaccine in high-income countries [@boyon2020ipsos; @Malik2020; @kreps2020factors; @shekhar2021covid; @dror2020vaccine; @fisher2020attitudes], and some studies have included middle-income countries [@lazarus2020nature; @wouters2021challenges]. Less is known, however, about vaccine acceptance in low-income countries where large-scale vaccination has yet to begin. Understanding the drivers of COVID-19 vaccine acceptance is of global concern, since a lag in vaccination in any country may result in the emergence and spread of new variants that can overcome immunity conferred by vaccines and prior disease [@ong2021lack; @lancettask].

Our study complements the emerging global picture of COVID-19 vaccine acceptance by focusing primarily on lower income countries. We construct a sample of low- and middle-income countries (LMICs) with wide geographic coverage across Africa, Asia and Latin America. We move beyond documenting vaccine acceptance rates to collect and analyze data on the reasons for acceptance and hesitancy, which is critical for informing the design of effective vaccine distribution and messaging. A summary of the main findings, limitations and implications of the study is shown in Table \ref{tab:summary}.

Acceptance of childhood vaccination for common diseases---such as measles (MCV), Bacille Calmette-Guérin (BCG) and diphtheria, tetanus, and pertussis (DTP)---is generally high in LMICs, providing grounds for optimism about the prospects for COVID-19 vaccine uptake. Table \ref{tab:otherv} summarizes general vaccine acceptance and coverage rates of childhood vaccines in 2018, prior to the current pandemic, for the countries included in our study. Agreement on the importance of childhood vaccinations is markedly higher in the LMICs we study compared to Russia and the United States. Still, existing studies on COVID-19 vaccine acceptance document substantial variation, both across and within countries, including in settings with high acceptance of other vaccinations [@defigueiredo2020lancet; @lazarus2020nature; @wouters2021challenges].

Existing literature cites concern about COVID-19 vaccine safety, including the rapid pace of vaccine development, as a primary reason for hesitancy in higher-income settings [@boyon2020ipsos; @wouters2021challenges]. Other reasons may feature more prominently in LMICs. For instance, reported COVID-19 mortality rates have been consistently lower in most LMICs relative to higher income countries [@biccard2021patient; @rice2021variation; @maeda2021puzzle]. If individuals feel the risk of disease is less severe, they may be less willing to accept any perceived risks of vaccination [@brewer2007meta]. Previous studies of healthcare utilization in LMICs have also highlighted factors such as negative perceptions of healthcare quality [@christensen2020building], negative historical experiences involving foreign actors  [@Lowes2018; @martinez2021vaccines], weak support from traditional leaders [@Jegede2007], and mistrust in government [@BLAIR201789] as barriers to uptake, which could apply to COVID-19 vaccination as well.

To promote vaccination against COVID-19, we need to know whether people are willing to take COVID-19 vaccines, the reasons why they are willing or unwilling to do so, and the most trusted sources of information in their decision-making. Our study investigates these questions using a common set of survey items deployed across 13 studies in Africa, South Asia, and Latin America: seven in low-income countries (Burkina Faso, Mozambique, Rwanda, Sierra Leone, Uganda), five in lower-middle-income countries (India, Nepal, Nigeria, Pakistan), and one in an upper-middle-income country (Colombia). We compare these findings to those from two countries at the forefront of vaccine research and development, Russia and the United States.

To select studies to include in our sample, we conducted an internal search within Innovations for Poverty Action (IPA), the International Growth Center (IGC), and the Berlin Social Science Center (WZB) for projects with plans to collect survey data in the second half of 2020. Study PIs agreed to include a set of common questions about COVID-19 vaccine attitudes. This strategy was guided by the need to collect information quickly and cost-effectively using a survey modality (phone) that was both safe, given pandemic conditions, and appropriate for contexts with limited internet coverage. The final set of samples included in our study therefore reflects populations that fall under the current research priorities at IGC, IPA and WZB and, in the case of IPA and IGC, donors that prioritize working in the Global South.

@@

Our main results are shown in Figure \@ref(fig:mainfigure1) and reproduced as \hyperref[tab:maintabledis]{Supplementary Table 1}. The first column provides overall acceptance rates in each study, while the remaining columns disaggregate acceptance by respondent characteristics. The “All LMICs” row reports averages for the LMIC samples included in our study and excludes Russia and the USA. The “LMICs National Samples” row reports averages for just the LMIC samples with national-level geographic coverage.

The average acceptance rate across the full set of LMIC studies is `r ans_mean`% (95% CI `r ans_mean_low`--`r ans_mean_high`%), with a median of `r ans_median`%, a range of `r ans_max - ans_min`% and an interquartile range of `r ans_iqr`%. Our estimate of the between-study standard deviation, $\tau$, using a random effects meta-analysis model is `r  round(sqrt(mp_tau), 3)` which represents only `r  round(100*mp_ratio, 1)`% of our estimate of the average acceptance across LMIC studies.

The acceptance rate in every LMIC sample is higher than in the USA (`r usa_ans %>% pull(estimate)`%, CI `r usa_ans %>% pull(conf.low)`--`r usa_ans %>% pull(conf.high)`%) and Russia (`r rus_ans %>% pull(estimate)`%, CI `r rus_ans %>% pull(conf.low)`--`r rus_ans %>% pull(conf.high)`%). Reported acceptance is lowest in `r bottom_all$group[1]` (`r bottom_all$estimate[1]`%, CI `r bottom_all$conf.low[1]`--`r bottom_all$conf.high[1]`%) and `r bottom_all$group[2]`  (`r bottom_all$estimate[2]`%, CI `r bottom_all$conf.low[2]`--`r bottom_all$conf.high[2]`%). Pakistan’s relatively low acceptance rate may be linked to negative historical experiences with foreign-led vaccination campaigns [@ali2019polio; @robbins2012cia; @martinez2021vaccines]. This hesitancy may be particularly problematic given the magnitude of the second wave in neighboring India and acceleration of cases across South Asia that threaten to overwhelm health infrastructure. The relatively low acceptance rate in Burkina Faso may reflect general vaccine hesitancy. As shown in Table \ref{tab:otherv}, fewer people believe that vaccines in general are safe in Burkina Faso than in any other country included in our study except Russia.

We find limited evidence of variation across demographic subgroups in our aggregate analysis of LMIC samples, as shown in \hyperref[tab:dmeans]{Supplementary Table 7}. Women are generally less willing to accept the vaccine than men (average difference about `r dif_gender` points, significant at $p<.01$). Respondents under age 25 and less educated respondents are marginally more willing to take the vaccine compared to older and more educated respondents, respectively, but these differences are not statistically significant.

Table \hyperref[tab:countrydiff]{Supplementary Table 9} provides results disaggregated by demographic subgroups for individual studies. The average gender differences in the aggregate LMIC analysis are driven by the Burkina Faso, Mozambique, Pakistan 1, Rwanda and Sierra Leone 1 samples. However, these gender differences in acceptance are less than 10 percentage points in each of these samples, in contrast to larger gender gaps in acceptance we observe in the USA (17%) and Russia (16%).

Less-educated respondents expressed significantly higher acceptance in the Burkina Faso, Rwanda, Sierra Leone 1, and Uganda 2 samples, which represent the majority of  studies from sub-Saharan Africa. Notably, we observe the opposite pattern in the India, Pakistan 1, and Pakistan 2 samples. In all three studies, acceptance is greater among more educated respondents, although this difference is not statistically significant in the India sample. Education is also a positive and significant predictor of acceptance in the USA.

We find mixed evidence across studies with respect to the relationship between age and COVID-19 vaccine acceptance. In India and Nigeria, respondents younger than 25 years old are significantly less willing to take the vaccine relative to adults 25-54 years old, while in Mozambique, Pakistan 1, and Rwanda, those under 25 are significantly more willing. In Mozambique and Rwanda, respondents under 25 are also significantly more accepting compared to those 55 and over; however, the difference between these age groups is not statistically significant in other LMIC samples. In the USA and Russia, older respondents are consistently more accepting than younger respondents.

To better understand the reasoning behind vaccine acceptance, we asked those who were willing to take the vaccine why they would take it. We summarize these results in \hyperref[tab:yes]{Supplementary Table 2}, with additional details in \hyperref[tab:yesall]{Supplementary Table 3}. The reason most commonly given for vaccine acceptance across samples is personal protection against COVID-19 infection. The average across the LMIC samples is `r yes_all$estimate`% (CI `r yes_all$conf.low`--`r yes_all$conf.high`%) with a median of  `r yes_ %>% pull(estimate) %>% median()`% and a range of `r yes_ %>% pull(estimate) %>% max() - yes_ %>% pull(estimate) %>% min()`%. In every individual study, including the USA (`r yes_usa$estimate`%, CI `r yes_usa$conf.low`--`r yes_usa$conf.high`%) and Russia (`r yes_rus$estimate`%, CI `r yes_rus$conf.low`--`r yes_rus$conf.high`%), this ranks as the most cited reason. In distant second place in the aggregate LMIC analysis is family protection, with an average of `r yes_all_2$estimate`% (CI `r yes_all_2$conf.low`--`r yes_all_2$conf.high`%), a median of `r yes_2 %>% pull(estimate) %>% median()`% and a range of `r yes_2 %>% pull(estimate) %>% max() - yes_2 %>% pull(estimate) %>% min()`%. In comparison to protecting oneself and one’s family,  protecting one’s community does not feature prominently among stated reasons for acceptance. These reasons do not vary substantially by age group, as shown in \hyperref[tab:yes1]{Supplementary Table 4}.

Figure \@ref(fig:fig2paper) summarizes the reasons given by respondents who said they were not willing to take a COVID-19 vaccine. Concern about side effects is the most frequently expressed reason for reluctance in our LMIC samples. This concern is particularly evident among samples from sub-Saharan Africa. In studies  `r side_top$group[1]` (`r side_top$estimate[1]`%, CI `r side_top$conf.low[1]`--`r side_top$conf.high[1]`%), `r side_top$group[2]`  (`r side_top$estimate[2]`%, CI `r side_top$conf.low[2]`--`r side_top$conf.high[2]`%), `r side_top$group[3]` (`r side_top$estimate[3]`%, CI `r side_top$conf.low[3]`--`r side_top$conf.high[3]`%) and `r side_top$group[4]` (`r side_top$estimate[4]`%, CI `r side_top$conf.low[4]`--`r side_top$conf.high[4]`%), more than half of respondents unwilling to take the vaccine cited worries about side effects.  Respondents in Russia (`r side_rus$estimate`%, CI `r side_rus$conf.low`--`r side_rus$conf.high`%) and even more in the USA (`r side_usa$estimate`%, CI `r side_usa$conf.low`--`r side_usa$conf.high`%), frequently report this same concern.

Study samples `r effective_top$group[1]` (`r effective_top$estimate[1]`%, CI `r effective_top$conf.low[1]`--`r effective_top$conf.high[1]`%), `r effective_top$group[2]`  (`r effective_top$estimate[2]`%, CI `r effective_top$conf.low[2]`--`r effective_top$conf.high[2]`%) and `r effective_top$group[3]` (`r effective_top$estimate[3]`%, CI `r effective_top$conf.low[3]`--`r effective_top$conf.high[3]`%) show relatively high levels of skepticism about vaccine effectiveness among hesitant respondents. This is also true in Russia (`r effective_rus$estimate`%, CI `r effective_rus$conf.low`--`r effective_rus$conf.high`%) and the USA (`r effective_usa$estimate`%, CI `r effective_usa$conf.low`--`r effective_usa$conf.high`%). In addition, some hesitant respondents cite lack of concern about COVID-19 infection as a reason not to be vaccinated. This answer is particularly common in the `r ill_top$group[1]` (`r ill_top$estimate[1]`%, CI `r ill_top$conf.low[1]`--`r ill_top$conf.high[1]`%), `r ill_top$group[2]`  (`r ill_top$estimate[2]`%, CI `r ill_top$conf.low[2]`--`r ill_top$conf.high[2]`%) and `r ill_top$group[3]` (`r ill_top$estimate[3]`%, CI `r ill_top$conf.low[3]`--`r ill_top$conf.high[3]`%) studies.

In Figure \@ref(fig:fig3paper) we report respondents’ most trusted source of guidance when deciding whether to take a COVID-19 vaccine. Results from Figure \@ref(fig:fig3paper) are reproduced in \hyperref[tab:trust]{Supplementary Table 6}, while \hyperref[tab:q4]{Supplementary Table 13} presents a complete description of response recoding from individual studies.

We find striking consistency across studies. In all samples except Rwanda, including those from Russia and the USA, respondents identify the health system as the most trustworthy source to help them decide whether to take the COVID-19 vaccine. The average across LMIC samples is  `r trust_all$estimate`% (CI `r trust_all$conf.low`--`r trust_all$conf.high`%), with a median of  `r trust_%>% pull(estimate) %>% median()`% and a range of `r trust_%>% pull(estimate) %>% max() - trust_%>% pull(estimate) %>% min()`%. Respondents in `r trust_top$group[1]` (`r trust_top$estimate[1]`%, CI `r trust_top$conf.low[1]`--`r trust_top$conf.high[1]`%), `r trust_top$group[2]`  (`r trust_top$estimate[2]`%, CI `r trust_top$conf.low[2]`--`r trust_top$conf.high[2]`%) and `r trust_top$group[3]` (`r trust_top$estimate[3]`%, CI `r trust_top$conf.low[3]`--`r trust_top$conf.high[3]`%) cited  health workers most often. Sierra Leone has the highest trust in health workers and the Ministry of Health, potentially reflecting investments in public health following the 2014-2015 Ebola epidemic [@deserrano].

In  `r trust_fam$group[1]` (`r trust_fam$estimate[1]`%, CI `r trust_fam$conf.low[1]`--`r trust_fam$conf.high[1]`%), `r trust_fam$group[2]`  (`r trust_fam$estimate[2]`%, CI `r trust_fam$conf.low[2]`--`r trust_fam$conf.high[2]`%), `r trust_fam$group[3]` (`r trust_fam$estimate[3]`%, CI `r trust_fam$conf.low[3]`--`r trust_fam$conf.high[3]`%) and `r trust_fam$group[4]` (`r trust_fam$estimate[4]`%, CI `r trust_fam$conf.low[4]`--`r trust_fam$conf.high[4]`%), the next most-cited sources are family and friends. Across the pooled samples, women are 3 percentage points more likely to rely on family and friends than male respondents, though this difference is not statistically significant (\hyperref[fig:genderhist]{Supplementary Figure 1}).
